Exercising choice over feedback schedules during practice is not advantageous for motor learning

### 2022-08-14 docs/revision-1/r1_response-to-reviewers.Rmd

---
title       : "Exercising choice over feedback schedules during practice is not advantageous for motor learning"
authors     : "Laura St. Germain, Brad McKay, Andrew Poskus, Allison Williams, Olena Leshchyshen, Sherry Feldman, Joshua G.A. Cashaback & Michael J. Carter"
journal     : "Psychonomic Bulletin & Review"
manuscript  : "PBR-BR-22-135"

bibliography      : "../references.bib"

class       : "final" # or final which removes all progress labels
output      : papaja::revision_letter_pdf
# output:
#   papaja::revision_letter_pdf:
#     latex_engine: xelatex

header-includes:
  - \usepackage[sfdefault,condensed]{roboto}
  - \usepackage[T1]{fontenc}
---

Melody Wiseheart, PhD  
Action Editor, _`r rmarkdown::metadata$journal`_

Dear Dr. Wiseheart,

Thank you for overseeing the review of our manuscript for publication in _`r rmarkdown::metadata$journal`_. We appreciate the positive and helpful comments from the Reviewers. Below we provide our point-by-point response (in normal font) to the Reviewers' comments (in ***bold-italic***). All substantial changes in the revised manuscript are included in this response letter (in a text box) and appear in \textcolor{blue}{blue} text in the revised manuscript.

Based on comments from all of the Reviewers, we have made our rationale for Experiment 2 more clear in the Introduction. Although the Reviewers made suggestions ranging from combining the two experiments to removing Experiment 2 altogether, we elected to keep it as Experiment 2. Doing so ensures transparency in our reporting practices as our decision to conduct Experiment 2 was made after we had analyzed the data for Experiment 1. Additionally, it is important to not censor null results in this area of motor learning research given the recent meta-analysis on this topic revealed the supposed self-controlled learning advantage resulted from publication bias rather than being a real effect [@mckay2021]. Reviewer 3 made several suggestions for additional analyses; however, we did not include some of them in the revision. Our primary reason for this decision was that the analysis suggestions were about understanding the information-processing explanation of self-controlled learning advantages. Yet, a key discussion point and conclusion based on our experiments is that there is no self-controlled learning advantage; a conclusion that is consistent with the @mckay2021 meta-analysis and several recent well-powered (and often pre-registered) experiments. We contend contextualizing our results within this current information negates the usefulness of additional analyses related to the information-processing (or motivational) perspective. We elaborate on these decisions in the point-by-point reply.

I would like to draw your attention to a change in the revised submission regarding our Supplementary documents. In our revision, the sensitivity analyses are now Supplementary A (were B in the original submission) and the error estimation data are now Supplementary B (were A in the original submission). To incorporate suggestions from the Reviewers, we have added Supplementary C which contains information related to the equivalence analyses on the psychological data. Thank you again for considering our work and overseeing the review process.

Sincerely,

Michael J. Carter

\Done

# Reviewer \#1

\RC{\emph{The authors present a well-designed investigation into self-control of feedback effects on motor skill learning. With N=228 participants assigned to self-controlled or yoked groups (crossed with feedback of differing quality: binary, graded, continuous), this is the largest study to date on the topic that I am aware of, and well-powered to test the a priori hypotheses. On the whole, I think the study is methodologically sound and of great interest to the motor learning community. That said, I do have some concerns about the structure of the manuscript, the use of equivalence tests, the interpretation of some of the data, which I detail below.}}

Thank you for the positive comments about our manuscript.

\Done

\RC{\emph{Validity of "Experiment 2". Given the lack of a learning effect in Experiment 2, I am not sure what this experiment really adds to the literature, nor do I think the self-control effect should be included in meta-analytic estimates moving forward (e.g., in this paper on p. 17, g=0.05, or in other future papers). In essence, the self-control of feedback is thought to moderate the learning effect, but with binary feedback there was no learning effect to begin with. Thus, it would be erroneous to include the null-effect of self-controlled feedback from that experiment when really there was simply no learning effect to moderate. I appreciate the authors reporting this experiment for completeness, but I think it's data should be removed from the equivalence testing section and the Discussion updated accordingly.}}

We agree that the lack of an improvement in performance during practice is unfortunate in Experiment 2. Despite this, we do think these results contribute to the literature. Our goal in the present project was to critically test between the information-processing and motivational (i.e., OPTIMAL theory) accounts of the (supposed) self-controlled learning advantage. After we finished analyzing the data from Experiment 1, it became clear that a second experiment was necessary where the feedback provided to participants would not generate an error signal, which occurs with both error and graded feedback (i.e., both of these provide information about how to correct the error). So binary feedback paired with choice (or no-choice) would be an even stronger test of the two perspectives than that in Experiment 1. From the information-processing perspective, you'd expect no performance advantage of having choice over feedback in practice, retention, or transfer because the information in the binary feedback will be poor at reducing uncertainties about performance outcomes. In contrast, from the motivational perspective (OPTIMAL theory) you'd expect a performance advantage of having choice over feedback in practice, retention, and transfer because the benefits arise from being given the opportunity to exercise choice. Thus, the lack of 1) an improvement in practice and 2) a learning effect in retention and transfer is consistent with the information-processing perspective. However, we do not interpret this as support for the information-processing perspective over the motivational view given the putative self-controlled learning advantage seems to be the result of publication bias rather than a real effect [@mckay2021]. We make note of this in a footnote on page 22. As such, any in-depth discussions about whether the information-processing or OPTIMAL theory is correct seem unwarranted. Based on your concerns, we have updated the Introduction (see below) to be more clear about our rationale and the value of Experiment 2.

> Introduction (Page 4 line 91 continued on page 5 lines 92-100):
>
> To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, we manipulated the amount of information participants in choice and yoked (i.e., no-choice) groups experienced with their feedback schedule during acqusition of a novel motor task. In Experiment 1, participants received error or graded feedback to assess how high and moderate levels of informational value impact the self-controlled learning advantage. Given both error and graded feedback provide salient information about how to correct one's behavior relative to the task goal (i.e., both generate an error signal), in Experiment 2 we provided participants with binary feedback. As binary feedback is devoid of information about the necessary change to improve one's behavior (i.e., does not generate an error signal), we could better isolate the motivational nature of choice to test between the two explanations for the self-controlled learning advantage.

Critically, we believe it is vital to report the combined estimate to combat publication bias, especially given the meta-analytic information that has begun to emerge revealing the putative self-controlled learning advantage is the result of publication bias rather than a real effect [@mckay2021]. We have kept the combined estimate and still focus on it in our Discussion; however, we now include in the Results section a separate estimate for each Experiment for the interested reader.

> Results (Page 19 Lines 349-352 continued on page 20 lines 353-355):
>
> Next, random effects meta-analyses were conducted on the retention test data^4^ to generate a summary point estimate and 90% confidence intervals (CI) with Experiments 1 and 2 combined and also separate. The overall estimated effect when combining both experiments} was *g* = .05 (favoring self-controlled) and 90% CI [-.12, .23]. The overall estimated effect for Experiment 1 was *g* = .03 (favoring self-controlled) and 90% CI [-.19, .25]. For Experiment 2, it was *g* = .09 (favoring self-controlled) and 90% CI [-.19, .37].

\Done

\RC{\emph{It is not clear to me why Experiment 2 is a separate experiment? This is the same task with a different type of feedback, right? Thus, I assume that this reflect the schedule of recruitment and all of the subjects in Experiment 2 were enrolled later. While not wrong to call it Experiment 2, I think it would be more parsimonious to have one experiment with 3 feedback types and then simply explain any cohort effects in the methods.}}

Yes, you are correct about the task, different type of feedback, and recruitment of participants for Experiment 2. We maintain referring to them as separate experiments as presenting it as a single experiment would pose some methodological problems. First, our decision to conduct Experiment 2 was conditional on observing the results of Experiment 1 (as mentioned in our above response), which would undermine the assumption of independence between observations. Second, random assignment to receive either error or graded feedback was done for participants in the self-controlled and yoked groups in Experiment 1. With the timing of the collection of these experiments, participants in Experiment 2 could not have been assigned to receive either error or graded feedback and Experiment 1 participants could not have been randomly assigned to receive binary feedback. Lastly, our *a priori* sample size calculation was not based on a design that included all groups in a single experiment.

\Done

\RC{\emph{Assignment of subjects to groups. From the authors' description, it sounds like ALL yoked participants were enrolled after ALL self-controlled participants in a given feedback condition. Albeit minor, this introduces a confound in the timing of the two groups. Something to consider for future studies would be a blocked assignment stratified by factors like gender age or handedness (e.g., the first right handed female to come into the lab gets self-control of a randomly assigned feedback type... the NEXT right handed female to come into the lab gets yoked to the same feedback schedule and type, etc). This would make the assignment order a bit more complicated, but would reduce the temporal disparity between the groups.}}

Yes, all participants in the self-controlled groups were collected before the yoked participants. We acknowledge in the original submission (and now Page 7 Lines 158-160 in the revised manuscript) that "this is typical in the self-controlled learning literature as the self-controlled participants' self-selected feedback schedules are required for providing feedback to the participants in the yoked (i.e., control) groups." We do agree that different assignment strategies could have been used; however, our decision was to follow the typical procedure for this area of research.

\Done

\RC{\emph{Lack of reporting for psychological variables. I might have missed it as there are many results presented, but could the authors report the (non-significant?) results of the self-control manipulation? The authors currently report the significant effects of Time and Feedback, but the effect of Self Control seems more theoretically relevant although non-significant.}}

We now report the effect of Choice in all analyses in the Results section. The effect of Choice was always non-significant.

\Done

\RC{\emph{Lack of equivalence tests for motivational variables. In the same way it is theoretically relevant to do equivalence tests at retention, it seems like it would be important to do equivalence tests for the different motivational variables as well?}}

We agree these would be theoretically relevant and have included them in Supplementary C.

\Done

# Review \#2

\RC{\emph{This manuscript reports a study that attempted to test different causal explanations for the self-controlled learning advantage. The explanations could not be examined because the experiment failed to find a self-controlled learning advantage. The study adds to a growing body of evidence which has failed to replicate the self-controlled learning advantage in the motor skills domain. Given the rigor with which the study was conducted and analyzed, the findings make an important contribution to the literature.}

\emph{I do not have any major concerns with the manuscript, though I may have called in to question the decision to limit practice to 72 trials had a self-controlled learning advantage been discovered. The reason is that Magill and Wood (1986) did not find a benefit of quantitative (error) versus qualitative (graded) feedback on a multi-movement timing task until the second half of practice. During the first half, learners derived the same benefit from quantitative and qualitative feedback. Consequently, it is not clear whether the amount of practice in the current experiment would have provided a fair test of the information processing account of the putative self-controlled learning advantage given that the informational value of quantitative feedback seems to increase with practice. Of course, that point is moot given the current findings.}}

Thank you for the encouraging and positive comments about our manuscript. The amount of practice is an interesting point given past motor learning research looking at this variable. We discuss the Magill and Wood paper below with another comment of yours.

\RC{\emph{In the abstract, the task is described as a rapid reaching task with spatial and temporal timing goals. I would not describe this task as a reaching task. I think it would be more appropriate to simply say “… while learning a planar upper limb movement with spatial and timing goals.” Note that it is referred to as a rapid aiming task on page 4, l. 25.}}

We have removed all reference to the task as a reaching task and now describe it as a "novel motor task" in the abstract.

\Done

\RC{\emph{I recommend stating that error and graded feedback have also been referred to as quantitative and qualitative feedback (e.g., Magill \& Wood, 1986) when you introduce error and graded feedback.}}

We have incorporated your suggestion as a footnote in the revised manuscript when we introduce error and graded feedback on page 3.

\Done

\RC{\emph{The description of the sample size estimation on page 5, l. 23 specifies that “groups = 2.” Shouldn’t the number of groups be 4?}}

Thank you for catching this typo. We have corrected it to say groups = 4 on page 6.

\Done

\RC{\emph{Comparisons were done to determine if error estimation improved with practice and whether the groups differed in terms of error estimation abilities at different stages of learning. However, it would be interesting to report if the participants who were asked to estimate during the pretest performed better in acquisition, retention and transfer than the participants who were not asked to estimate during the pretest. I realize this comparison was not an original purpose of the experiment, however, reporting this information would add value in my opinion.}}

We have added this analysis to Supplementary B and there were no significant main effects or interactions.

\Done

\RC{\emph{I understand the rationale for restricting the participants in the choice groups to 24 trials, however, it would be prudent to raise the possibility in the discussion that limiting choice in this way may have attenuated the self-controlled learning advantage.}}

We do not think this restriction contributed to our lack of a self-controlled learning advantage. Based on a recent meta-analysis [@mckay2021], we believe we did not find a self-controlled learning advantage because it either does not exist or is an extremely small effect. However, to your point earlier self-controlled feedback research actually found that limiting the number of choice opportunities does not negatively impact learning [@patterson2011] and can even lead to superior learning [@hansen2011]. Based on the conclusions of the self-controlled learning meta-analysis, we elected to not discuss this possibility and instead maintain the focus on our results being consistent with the meta-analytic estimates and conclusions of @mckay2021 regarding the self-controlled learning advantage being either a non-real effect or an extremely small effect.

\Done

\RC{\emph{I am not certain the findings from Experiment 2 offer much insight into self-controlled learning given that learners were not able to learn anything with this type of feedback. However, for the sake of thoroughness, Experiment 2 deserves to be described in the manuscript. I recommend reporting the number of “hits” in Experiment 2 to give readers a sense of the number of occasions learners received some sort of positive reinforcement. I think it would also be helpful to report the number of hits in Experiment 1 to potentially bolster the argument for the proposed disassociation between the informational and motivational impacts of knowledge of results.}}

Based on your comment (and a comment by Reviewer 1), we have updated the introduction to more clearly describe our rationale for and the value of Experiment 2. The lack of a performance improvement in practice is consistent with the information-processing perspective as the binary feedback would not be effective at reducing uncertainties about performance outcomes. As such, under such conditions you would expect the results we saw. However, the lack of improvement is inconsistent with a motivational perspective wherein having choice itself is the cause for any performance and learning benefits. We do not interpret this as support for the information-processing perspective over the motivational view given the putative self-controlled learning advantage seems to be the result of publication bias rather than a real effect [@mckay2021]. We make note of this in a footnote on page 22.

> Introduction (Page 4 line 91 continued on page 5 lines 92-100):
>
> To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, we manipulated the amount of information participants in choice and yoked (i.e., no-choice) groups experienced with their feedback schedule during acqusition of a novel motor task. In Experiment 1, participants received error or graded feedback to assess how high and moderate levels of informational value impact the self-controlled learning advantage. Given both error and graded feedback provide salient information about how to correct one's behavior relative to the task goal (i.e., both generate an error signal), in Experiment 2 we provided participants with binary feedback. As binary feedback is devoid of information about the necessary change to improve one's behavior (i.e., does not generate an error signal), we could better isolate the motivational nature of choice to test between the two explanations for the self-controlled learning advantage.

Thank you for the suggestion of including the number of hits in both experiments. We have included this information in Table 2 (page 18) in the revised manuscript and we also make reference to it in the Discussion.

> Discussion (Page 23 Lines 447-450):
>
> The number of "hits" for the spatial and timing goals were both quite low for all groups. Although this may have also impacted perceptions of competence, the relatively low "hit" rate did not seem to differentially impact intrinsic motivation as self-reported levels were quite similar for all groups.

\Done

\RC{\emph{Were the questions from the Intrinsic Motivation Inventory modified in any way and were the perceived autonomy questions exactly the same as those used earlier?}}

We modified the questions for the IMI subscales so it referenced "the motor task" when participants were asked the various questions. An example for the perceived competence subscale is: "I think I am pretty good at this motor task". An example for the interest/enjoyment is: "This motor task was fun to do". The perceived autonomy questions (e.g., "There is not much opportunity for my to exercise choices during my practice session") were the same across both experiments and used in previous work [e.g., @carter2017b; @stgermain2022]. We have included a copy of all the questions used for perceived competence, intrinsic motivation, and perceived autonomy in the public repository for this experiment in a directory called "materials". We make a specific note of the availability of the questionnaires in a footnote on page 11.

\Done

\RC{\emph{Do you have any explanation for why the graded feedback group continued to improve during practice after the error feedback group had plateaued? This is a curious finding, as one would expect just the opposite based on Magill and Wood (1986).}}

It was an interesting result that @magill1986 did not find a difference in feedback types until the second half of practice with 100 trials; although there are some considerable differences between their experiment and ours. First, they used a motor task that required six movements and varying timing goals compared to our single movement with a timing and spatial goal. So there could be some differences due to complexity or difficulty of the to-be-learned motor task. Second, Magill and Wood report that they did not find an interaction between Block and Feedback type until they strayed from their *a priori* analysis plan and excluded a block after observing their data. This interaction was also only found in one of their three performance outcome variables. For these reasons, we think their result must be interpreted cautiously. However, we do think it would be of value for future studies to further probe this potential interaction between amount of practice and different characteristics of the feedback provided to the learners. In terms of the significant Feedback x Block interaction we found for the timing goal in Experiment 1, the improvements in the Graded feedback groups after block 2 were quite minimal. In Figure 2B, all four groups are very similar from blocks 4 to 6. A potential explanation for this could be that the high informational value of the error feedback could have led to those participants making short-term maladaptive corrections to their responses [@salmoni1984].

\Done

# Reviewer \#3

\RC{\emph{This paper reports two experiments (total n=228) on how exercising choice influences motor learning of a reaching task with spatial and timing goals. Participants practiced either with choice over feedback requests or under no-choice, yoked feedback conditions. In Experiment 1, two forms of knowledge of results feedback were tested (error feedback or graded feedback) and in Experiment 2, binary feedback was provided. Behavioral performance, error estimation, and psychosocial data were collected. Whereas the error and graded feedback conditions supported skill acquisition, the binary feedback did not yield significant improvements from pre-testing to retention. Across both experiments, there was no evidence that the opportunity to exert self-control enhanced learning compared to the no-choice groups.}

### 2022-08-14 docs/revision-1/r1_manuscript.Rmd

  \noindent \textbf{Corresponding authors:} Michael J. Carter (michaelcarter@mcmaster.ca) or Laura St. Germain (stgerml@mcmaster.ca)

abstract: |
  The idea that there is a self-controlled learning advantage, where individuals demonstrate improved motor learning after exercising choice over an aspect of practice compared to no-choice groups, has different causal explanations according to the OPTIMAL theory or an information-processing perspective. Within OPTIMAL theory, giving learners choice is considered an autonomy-supportive manipulation that enhances expectations for success and intrinsic motivation. In the information-processing view, choice allows learners to engage in performance-dependent strategies that reduce uncertainty about task outcomes. To disentangle these potential explanations, we provided participants in choice and yoked groups with error or graded feedback (Experiment 1) and binary feedback (Experiment 2) while learning a rapid reaching task with spatial and timing goals. Across both experiments (N = 228 participants), we did not find evidence to support a self-controlled learning advantage. Exercising choice during practice did not increase perceptions of autonomy, competence, or intrinsic motivation, nor did it lead to more accurate error estimation skills. Both error and graded feedback facilitated skill acquisition and learning, whereas no improvements from pre-test performance were found with binary feedback. Finally, the impact of graded and binary feedback on perceived competence highlights a potential dissociation of motivational and informational roles of feedback. Although our results regarding self-controlled practice conditions are difficult to reconcile with either the OPTIMAL theory or the information-processing perspective, they are consistent with a growing body of evidence that strongly suggests self-controlled conditions are not an effective approach to enhance motor performance and learning.

keywords          : "Self-controlled, Knowledge of results, OPTIMAL theory, Retention"
wordcount         : "2978 for Introduction, Results, and Discussion"

@@

header-includes   :
  - \usepackage{pdflscape}
  - \usepackage{setspace}
  - \pagewiselinenumbers
  - \raggedbottom

@@

The underlying source of errors in skilled actions are often ambiguous and difficult to assign as the learner must rely on noisy and delayed sensory information. Feedback from an external source, such as a coach or computer display, can facilitate or augment this process [@sigrist2013]. Knowledge of results feedback [@salmoni1984] can provide varying amounts of information to learners depending on its characteristics. Error feedback provides precise information about the magnitude and direction of the error (e.g., -42 cm), graded feedback provides coarse information about either the magnitude or direction of the error (e.g., "too far"), and binary feedback indicates only success or failure information (e.g., "miss") [@luft2014]. When to provide this feedback is often decided by an external agent; however, this feedback decision can also be made by the learner, a form of self-controlled learning. These self-controlled feedback schedules have typically enhanced motor skill learning compared to yoked feedback schedules, wherein learners experience the feedback schedule created by a self-controlled counterpart, but without any choice [see @sanli2013; @stemarie2019 for reviews].

@@

To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, we provided participants in choice and yoked (i.e., no-choice) groups with error or graded feedback (Experiment 1) and binary feedback (Experiment 2) during the acquisition of a rapid aiming task. Motor learning was assessed using delayed (~24 hours) retention and transfer tests. If the OPTIMAL theory is correct, we hypothesized that the characteristics of one's feedback schedule *would not matter* for the self-controlled learning advantage as this advantage arises from the opportunity for choice--a common feature of all choice groups. Thus, we predicted all choice groups would demonstrate superior performance and learning compared to the yoked groups. Alternatively, if the information-processing account is correct, we hypothesized that the characteristics of one's feedback schedule *would matter* for the self-controlled learning advantage as feedback with greater informational value would be more effective for reducing uncertainties about movement outcomes. Thus, we predicted that choice over an error feedback schedule would be the most effective pairing for performance and learning. We also included self-report measures of perceptions of autonomy, competence, and intrinsic motivation, and assessments of error estimation abilities to respectively test auxiliary assumptions of the OPTIMAL theory and information-processing explanations.

@@

One hundred and fifty-two right-handed [@oldfield1971], healthy adults participated in Experiment 1 ($M_{age}$ = 20.64 years, $SD_{age}$ = 2.45, 88 females). Sample size was determined from an *a-priori* power analysis using the *ANOVA: fixed effects, main effects and interactions* option in G*Power [@faul2009] with the following parameters: $\alpha$ = 0.05, $\beta$ = .20, $f$ = 0.23, numerator = 1, and groups = 2. This revealed a required sample of 151 participants. Our chosen effect size was based on a meta-analytic estimate ($f$ = .32) by @mckay2014; however, we used a more conservative estimate given the uncertainty of how choice would interact with our feedback characteristic manipulation. Participants were compensated $15 CAD or with course-credit for their time. All participants gave written informed consent and the experiment was approved by McMaster University’s Research Ethics Board.

@@

Data collection consisted of two sessions separated by approximately 24 hours.[^1] Session one included a pre-test (12 trials) and an acquisition phase (72 trials). Session two included the delayed retention (12 trials) and transfer (12 trials) tests. No feedback about motor performance was provided in pre-test, retention, or transfer. Prior to the pre-test, all participants received instructions about the task and its associated spatial and timing goals. Additionally, half of the participants in each group were randomly selected to estimate their performance on the spatial and timing goals after each trial in the pre-test. Only a subset of participants were asked to estimate their performance in pre-test to mitigate the potential that doing so would prompt participants to adopt this strategy during the experiment as error estimation has been suggested [e.g., @chiviacowsky2005] to be adopted spontaneously by participants controlling their feedback schedule. However, asking participants to estimate their performance during pre-test is necessary to be able to assess how this skill develops as a function of one's practice condition.

[^1]: Six participants (three Choice+Error-Feedback and three Choice+Graded-Feedback) had their second session completed approximately 48 hours later because a snowstorm closed the University.

@@

A typical acquisition trial (see Figure \@ref(fig:fig1)) began with the current trial number displayed (500 ms), followed by a visual "Get Ready!" and a visual go-signal (800 ms apart). Participants were free to begin their movement when ready following the visual go-signal (i.e., green circle) as this was not a reaction time task. The computer screen was blank while participants made their movement. When participants returned to the starting position, a red circle was displayed on the monitor. Following a 2000 ms feedback delay interval, the feedback decision prompt was presented for the self-controlled groups. If feedback was not selected, a blank screen was displayed for 3000 ms. If feedback was selected (or imposed on the yoked groups), it was also displayed for 3000 ms.

Before the retention and transfer tests, participants were reminded about the task and its associated goals. All participants were asked to estimate their performance after each trial in retention and transfer. After the pre-test, trials 12 and 72 in acquisition, and before the delayed retention test, participants answered a series of questions pertaining to perceived competence, task interest and enjoyment, and perceived autonomy. The perceived competence and task interest and enjoyment questions were from the Intrinsic Motivation Inventory [@mcauley1989; @ryan1982] and the perceived autonomy questions were used in earlier work [@barros2019; @carter2017b]. Cronbach's alpha values for each questionnaire at each time point are reported in Table \@ref(tab:table1).

@@

(ref:fig1-caption) \small \onehalfspacing \textbf{Overview of a typical acquisition trial for the choice groups.} The sequence of events a participant in the choice groups experienced during the acquisition phase. Trials began by informing participants the trial number (500 ms) they were on in acquisition. Shortly after, the text "Get Ready!" appeared on the screen and 800 ms later a visual go-signal was presented in the form of a green circle in the center of the screen. Participants began their movement when ready after seeing the visual go-signal as we were not interested in reaction time. While participants completed their rapid out-and-back movement, the computer screen was blank. Upon returning to the starting position, a red circle appeared in the center of the screen. A 2000 ms feedback delay interval was used and this interval was followed by the feedback prompt. The feedback prompt also displayed an updated counter representing the number of feedback trials they had left. If the number of remaining feedback trials matched the number of acquisition trials left, these trials automatically defaulted to feedback trials. On trials where feedback was not requested, a blank screen **(A)** was shown for 3000 ms. When feedback was selected, feedback was provided for both the spatial and timing goals according to their experimental group. The error feedback group **(B)** saw their constant error, the graded feedback group **(C)** saw either "too far" or "too short" for the spatial goal and "too fast" or "too slow" for the timing goal, and the binary feedback group **(D)** saw either "hit" or "miss" for the task goals. The sequence of events was the same for the yoked groups with the exception they did not see a feedback prompt. The sequence of events was similar in pre-test, retention, and transfer except all trials were no-feedback trials.

@@

Model diagnostics of total error for the spatial and timing goals revealed skewed distributions. We therefore conducted sensitivity analyses using the shift function, which is a robust statistical method well-suited for skewed distributions [@rousselet2020; @wilcox2021]. The results of these analyses (see **Supplementary B**) were consistent with those of the mixed ANOVAs, which we report below.

>>>>

  \noindent \textbf{Corresponding authors:} Laura St. Germain (stgerml@mcmaster.ca) or Michael J. Carter (cartem11@mcmaster; motorlab@mcmaster.ca)

abstract: |
  The idea that there is a self-controlled learning advantage, where individuals demonstrate improved motor learning after exercising choice over an aspect of practice compared to no-choice groups, has different causal explanations according to the OPTIMAL theory or an information-processing perspective. Within OPTIMAL theory, giving learners choice is considered an autonomy-supportive manipulation that enhances expectations for success and intrinsic motivation. In the information-processing view, choice allows learners to engage in performance-dependent strategies that reduce uncertainty about task outcomes. To disentangle these potential explanations, we provided participants in choice and yoked groups with error or graded feedback (Experiment 1) and binary feedback (Experiment 2) while learning a \textcolor{blue}{novel motor} task with spatial and timing goals. Across both experiments (N = 228 participants), we did not find evidence to support a self-controlled learning advantage. Exercising choice during practice did not increase perceptions of autonomy, competence, or intrinsic motivation, nor did it lead to more accurate error estimation skills. Both error and graded feedback facilitated skill acquisition and learning, whereas no improvements from pre-test performance were found with binary feedback. Finally, the impact of graded and binary feedback on perceived competence highlights a potential dissociation of motivational and informational roles of feedback. Although our results regarding self-controlled practice conditions are difficult to reconcile with either the OPTIMAL theory or the information-processing perspective, they are consistent with a growing body of evidence that strongly suggests self-controlled conditions are not an effective approach to enhance motor performance and learning.

keywords          : "Self-controlled, Knowledge of results, OPTIMAL theory, Retention"
#wordcount         : "2978 for Introduction, Results, and Discussion"

@@

header-includes   :
  - \usepackage{pdflscape}
  - \usepackage{setspace}
  #- \pagewiselinenumbers
  - \raggedbottom

@@

The underlying source of errors in skilled actions are often ambiguous and difficult to assign as the learner must rely on noisy and delayed sensory information. Feedback from an external source, such as a coach or computer display, can facilitate or augment this process [@sigrist2013]. Knowledge of results feedback [@salmoni1984] can provide varying amounts of information to learners depending on its characteristics. Error feedback provides precise information about the magnitude and direction of the error (e.g., -42 cm), graded feedback provides coarse information about either the magnitude or direction of the error (e.g., "too far"), and binary feedback indicates only success or failure information (e.g., "miss") [@luft2014].[^1] When to provide this feedback is often decided by an external agent; however, this feedback decision can also be made by the learner, a form of self-controlled learning. These self-controlled feedback schedules have typically enhanced motor skill learning compared to yoked feedback schedules, wherein learners experience the feedback schedule created by a self-controlled counterpart, but without any choice [see @sanli2013; @stemarie2019 for reviews].

[^1]: \textcolor{blue}{Others have referred to error feedback as quantitative feedback and graded feedback as qualitative feedback} [e.g., @magill1986]\textcolor{blue}{. We use the terminology error, graded, and binary feedback because graded and binary feedback are different forms of qualitative feedback.}

@@

To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, \textcolor{blue}{we manipulated the amount of information participants in choice and yoked (i.e., no-choice) groups experienced with their feedback schedule during acqusition of a novel motor task. In Experiment 1, participants received error or graded feedback to assess how high and moderate levels of informational value impact the self-controlled learning advantage. Given both error and graded feedback provide salient information about how to correct one's behavior relative to the task goal (i.e., both generate an error signal), in Experiment 2 we provided participants with binary feedback. As binary feedback is devoid of information about the necessary change to improve one's behavior (i.e., does not generate an error signal), we could better isolate the motivational nature of choice to test between the two explanations for the self-controlled learning advantage.} Motor learning was assessed using delayed (~24 hours) retention and transfer tests. If the OPTIMAL theory is correct, we hypothesized that the characteristics of one's feedback schedule *would not matter* for the self-controlled learning advantage as this advantage arises from the opportunity for choice--a common feature of all choice groups. Thus, we predicted all choice groups would demonstrate superior performance and learning compared to the yoked groups. Alternatively, if the information-processing account is correct, we hypothesized that the characteristics of one's feedback schedule *would matter* for the self-controlled learning advantage as feedback with greater informational value would be more effective for reducing uncertainties about movement outcomes. Thus, we predicted that choice over an error feedback schedule would be the most effective pairing for performance and learning. We also included self-report measures of perceptions of autonomy, competence, and intrinsic motivation, and assessments of error estimation abilities to respectively test auxiliary assumptions of the OPTIMAL theory and information-processing explanations.

@@

One hundred and fifty-two right-handed [@oldfield1971], healthy adults participated in Experiment 1 ($M_{age}$ = 20.64 years, $SD_{age}$ = 2.45, 88 females). Sample size was determined from an *a-priori* power analysis using the *ANOVA: fixed effects, main effects and interactions* option in G\*Power [@faul2009] with the following parameters: $\alpha$ = 0.05, $\beta$ = .20, $f$ = 0.23, numerator = 1, and groups = \textcolor{blue}{4}. This revealed a required sample of 151 participants. Our chosen effect size was based on a meta-analytic estimate (*f* = .32) by @mckay2014; however, we used a more conservative estimate given the uncertainty of how choice would interact with our feedback characteristic manipulation. Participants were compensated $15 CAD or with course-credit for their time. All participants gave written informed consent and the experiment was approved by McMaster University’s Research Ethics Board.

@@

Data collection consisted of two sessions separated by approximately 24 hours.[^2] Session one included a pre-test (12 trials) and an acquisition phase (72 trials). Session two included the delayed retention (12 trials) and transfer (12 trials) tests. No feedback about motor performance was provided in pre-test, retention, or transfer. Prior to the pre-test, all participants received instructions about the task and its associated spatial and timing goals. Additionally, half of the participants in each group were randomly selected to \textcolor{blue}{verbally} estimate their performance on the spatial and timing goals after each trial in the pre-test. Only a subset of participants were asked to estimate their performance in pre-test to mitigate the potential that doing so would prompt participants to adopt this strategy during the experiment as error estimation has been suggested [e.g., @chiviacowsky2005] to be adopted spontaneously by participants controlling their feedback schedule. However, asking participants to estimate their performance during pre-test is necessary to be able to assess how this skill develops as a function of one's practice condition.

[^2]: Six participants (three Choice+Error-Feedback and three Choice+Graded-Feedback) had their second session completed approximately 48 hours later because a snowstorm closed the University.

@@

A typical acquisition trial (see Figure \@ref(fig:fig1)) began with the current trial number displayed (500 ms), followed by a visual "Get Ready!" and a visual go-signal (800 ms apart). Participants were free to begin their movement when ready following the visual go-signal (i.e., green circle) as this was not a reaction time task. The computer screen was blank while participants made their movement. When participants returned to the starting position, a red circle was displayed on the monitor. Following a 2000 ms feedback delay interval, the feedback decision prompt was presented for the self-controlled groups. \textcolor{blue}{The number of remaining feedback trials was also displayed during this feedback delay interval.} If feedback was not selected, a blank screen was displayed for 3000 ms. If feedback was selected via \textcolor{blue}{verbal response} (or imposed on the yoked groups), it was also displayed for 3000 ms.

\clearpage

(ref:fig1-caption) \small \normalfont \onehalfspacing \textbf{Overview of a typical acquisition trial for the choice groups.} The sequence of events a participant in the choice groups experienced during the acquisition phase. Trials began by informing participants the trial number (500 ms) they were on in acquisition. Shortly after, the text "Get Ready!" appeared on the screen and 800 ms later a visual go-signal was presented in the form of a green circle in the center of the screen. Participants began their movement when ready after seeing the visual go-signal as we were not interested in reaction time. While participants completed their rapid out-and-back movement, the computer screen was blank. Upon returning to the starting position, a red circle appeared in the center of the screen. A 2000 ms feedback delay interval was used and this interval was followed by the feedback prompt. The feedback prompt also displayed an updated counter representing the number of feedback trials they had left. If the number of remaining feedback trials matched the number of acquisition trials left, these trials automatically defaulted to feedback trials. On trials where feedback was not requested, a blank screen **(A)** was shown for 3000 ms. When feedback was selected via \textcolor{blue}{verbal response}, feedback was provided for both the spatial and timing goals according to their experimental group. The error feedback group **(B)** saw their constant error, the graded feedback group **(C)** saw either "too far" or "too short" for the spatial goal and "too fast" or "too slow" for the timing goal, and the binary feedback group **(D)** saw either "hit" or "miss" for the task goals. The sequence of events was the same for the yoked groups with the exception they did not see a feedback prompt. The sequence of events was similar in pre-test, retention, and transfer except all trials were no-feedback trials.

\clearpage

Before the retention and transfer tests, participants were reminded about the task and its associated goals. All participants were asked to \textcolor{blue}{verbally} estimate their performance after each trial in retention and transfer. After the pre-test, trials 12 and 72 in acquisition, and before the delayed retention test, participants \textcolor{blue}{verbally} answered a series of questions pertaining to perceived competence, task interest and enjoyment, and perceived autonomy.[^3] The perceived competence and task interest and enjoyment questions were from the Intrinsic Motivation Inventory [@mcauley1989; @ryan1982] and the perceived autonomy questions were used in earlier work [@barros2019; @carter2017b; @stgermain2022]. Cronbach's alpha values for each questionnaire at each time point are reported in Table \@ref(tab:table1).

[^3]: \textcolor{blue}{The questionnaires can be found in the publicly available project repository in the materials directory.}

@@

@@
